---
title: "Project_use"
author: "Isabela Vieira"
date: "12/9/2020"
output:
  word_document: default
  pdf_document: default
  'word_document: default': default
---

#Re-coding some of the continuous variables in ACS (Levels and colnames resembles BRFSS levels)




```{r}
load("~/Documents/College/Fall 2020/Econometrics/acs2017_ny_data.RData")
BRFSS_USE <- readRDS("~/Documents/College/Fall 2020/Econometrics/Final Project/FINAL PROJECT/BRFSS_USE.rds")
attach(acs2017_ny)
attach(BRFSS_USE)

```

```{r, echo=FALSE}
#Income variable

library(forcats)

acs2017_ny$INCOME2[acs2017_ny$HHINCOME > 0 & acs2017_ny$HHINCOME < 10000] = "10000"
acs2017_ny$INCOME2[acs2017_ny$HHINCOME >= 10000 & acs2017_ny$HHINCOME < 15000] = "15000"
acs2017_ny$INCOME2[acs2017_ny$HHINCOME >= 15000 & acs2017_ny$HHINCOME < 20000] = "20000"
acs2017_ny$INCOME2[acs2017_ny$HHINCOME >= 20000 & acs2017_ny$HHINCOME < 25000] = "25000"
acs2017_ny$INCOME2[acs2017_ny$HHINCOME >= 25000 & acs2017_ny$HHINCOME < 35000] = "35000"
acs2017_ny$INCOME2[acs2017_ny$HHINCOME >= 35000 & acs2017_ny$HHINCOME < 50000] = "50000"
acs2017_ny$INCOME2[acs2017_ny$HHINCOME >= 50000 & acs2017_ny$HHINCOME < 75000] = "75000"
acs2017_ny$INCOME2[acs2017_ny$HHINCOME >= 75000] = "2030000"

acs2017_ny$INCOME2 = factor(acs2017_ny$INCOME2,levels=c("10000", "15000", "20000","25000","35000","50000","75000","2030000"))

#Age 
acs2017_ny$AGE[acs2017_ny$AGE >= 18 & acs2017_ny$AGE <= 24] = "18"
acs2017_ny$AGE[acs2017_ny$AGE >= 25 & acs2017_ny$AGE <= 29] = "25"
acs2017_ny$AGE[acs2017_ny$AGE >= 30 & acs2017_ny$AGE <= 34] = "30"
acs2017_ny$AGE[acs2017_ny$AGE >= 35 & acs2017_ny$AGE <= 39] = "35"
acs2017_ny$AGE[acs2017_ny$AGE >= 40 & acs2017_ny$AGE <= 44] = "40"
acs2017_ny$AGE[acs2017_ny$AGE >= 45 & acs2017_ny$AGE <= 49] = "45"
acs2017_ny$AGE[acs2017_ny$AGE >= 50 & acs2017_ny$AGE <= 54] = "50"
acs2017_ny$AGE[acs2017_ny$AGE >= 55 & acs2017_ny$AGE <= 59] = "55"
acs2017_ny$AGE[acs2017_ny$AGE >= 60 & acs2017_ny$AGE <= 64] = "60"
acs2017_ny$AGE[acs2017_ny$AGE >= 65 & acs2017_ny$AGE <= 69] = "65"
acs2017_ny$AGE[acs2017_ny$AGE >= 70 & acs2017_ny$AGE <= 74] = "70"
acs2017_ny$AGE[acs2017_ny$AGE >= 75 & acs2017_ny$AGE <= 79] = "75"
acs2017_ny$AGE[acs2017_ny$AGE >= 80] = "80"

acs2017_ny$AGE = factor(acs2017_ny$AGE,levels=c("18","25","30","35","40","45","50","55","60","65","70","75","80"))


#Ownership 

is.na(acs2017_ny$OWNERSHP) <- which(acs2017_ny$OWNERSHP == 0)
na.exclude(acs2017_ny$OWNERSHP)
acs2017_ny$RENTHOM1 <- as.factor(acs2017_ny$OWNERSHP)
levels(acs2017_ny$RENTHOM1) <- c("Owned","Rent")


#Marital Status 

acs2017_ny$MARST <- as.factor(acs2017_ny$MARST) #Have to change name in BRFSS
levels(acs2017_ny$MARST) <- c("married","married2","separated","divorced","widowed","never.married")
acs2017_ny$MARST <- fct_collapse(acs2017_ny$MARST, married = c("married","married2"))
fct_relevel(acs2017_ny$MARST , "divorced", "married", "never.married","separated", "widowed")

#Gender
acs2017_ny$SEXVAR <- as.factor(acs2017_ny$female)
levels(acs2017_ny$SEXVAR) <- c("Male","Female")

#Race

acs2017_ny$RACE <- as.factor(acs2017_ny$RACE) #Have to change name in BRFSS
levels(acs2017_ny$RACE) <- c("White","Black","American Indian/Alaskan Native", "Chinese","Japanese","Asian other","Other Race","Two major races","Three or more major races")
acs2017_ny$RACE <- fct_collapse(acs2017_ny$RACE, Asian = c("Chinese","Japanese","Asian other"))
acs2017_ny$RACE <- fct_collapse(acs2017_ny$RACE, Multiracial = c("Two major races","Three or more major races"))
fct_relevel(acs2017_ny$RACE, "American Indian/Alaskan Native", "Asian", "Black","Multiracial", "Other Race", "White")

#Hispanic

acs2017_ny$X_HISPANC <- as.factor(acs2017_ny$Hispanic)
levels(acs2017_ny$X_HISPANC) <- c("No","Yes")
fct_relevel(acs2017_ny$X_HISPANC, "Yes", "No")

#Health Coverage

acs2017_ny$HLTHPLN1 <- as.factor(acs2017_ny$HCOVANY)
levels(acs2017_ny$HLTHPLN1) <- c("No","Yes")


#In Labor force

is.na(acs2017_ny$LABFORCE) <- which(acs2017_ny$LABFORCE == 0)
na.exclude(acs2017_ny$LABFORCE)
acs2017_ny$LABFORCE <- as.factor(acs2017_ny$LABFORCE)
levels(acs2017_ny$LABFORCE) <- c("Not.in.LF","in.LF")

#Education

acs2017_ny$EDUC <- as.numeric(acs2017_ny$EDUC)
is.na(acs2017_ny$EDUC) <- which(acs2017_ny$EDUC == 00)
na.exclude(acs2017_ny$EDUC)
acs2017_ny$EDUCA <- as.factor(acs2017_ny$EDUC)
levels(acs2017_ny$EDUCA) <- c("n to 4","5 to 8","9","10","11","12", "College 1 year", "College 2 ", "College 3", "College 4 ", "College 5")
acs2017_ny$EDUCA <- fct_collapse(acs2017_ny$EDUCA, Elementary = c("n to 4","5 to 8"))
acs2017_ny$EDUCA <- fct_collapse(acs2017_ny$EDUCA, SomeHS = c("9","10","11"))
acs2017_ny$EDUCA <- fct_collapse(acs2017_ny$EDUCA, HS = c("12"))
acs2017_ny$EDUCA <- fct_collapse(acs2017_ny$EDUCA, SomeC = c("College 1 year", "College 2 ", "College 3"))
acs2017_ny$EDUCA <- fct_collapse(acs2017_ny$EDUCA, College = c("College 4 ", "College 5"))

acs2017_ny$in_NYC <- as.factor(acs2017_ny$in_NYC)
levels(acs2017_ny$in_NYC) <- c("No","Yes")

```


#Recoding for BRFSS
```{r, echo=FALSE}
#INCOME
is.na(BRFSS_USE$INCOME2) <- which(BRFSS_USE$INCOME2 == 99)
is.na(BRFSS_USE$INCOME2) <- which(BRFSS_USE$INCOME2 == 77)
na.exclude(BRFSS_USE$INCOME2)
BRFSS_USE$INCOME2 <- as.factor(BRFSS_USE$INCOME2)
levels(BRFSS_USE$INCOME2) <- c("0", "10000", "15000","20000","25000","35000","50000","75000","2030000")

#Age Variable 

is.na(BRFSS_USE$X_AGEG5YR) <- which(BRFSS_USE$X_AGEG5YR == 14)
na.exclude(BRFSS_USE$X_AGEG5YR)
BRFSS_USE$AGE <- as.factor(BRFSS_USE$X_AGEG5YR)
levels(BRFSS_USE$AGE) <- c("18","25","30","35","40","45","50","55","60","65","70","75","80")

#Gender 

BRFSS_USE$SEXVAR <- as.factor(BRFSS_USE$SEXVAR)
levels(BRFSS_USE$SEXVAR) <- c("Male","Female")


#Ownership 
is.na(BRFSS_USE$RENTHOM1) <- which(BRFSS_USE$RENTHOM1 == 7)
is.na(BRFSS_USE$RENTHOM1) <- which(BRFSS_USE$RENTHOM1 == 9)
is.na(BRFSS_USE$RENTHOM1) <- which(BRFSS_USE$RENTHOM1 == 3)
na.exclude(BRFSS_USE$RENTHOM1)
BRFSS_USE$RENTHOM1 <- as.factor(BRFSS_USE$RENTHOM1)
levels(BRFSS_USE$RENTHOM1) <- c("Owned","Rent")

#Marital Status 

is.na(BRFSS_USE$MARITAL) <- which(BRFSS_USE$MARITAL == 9)
na.exclude(BRFSS_USE$MARITAL)
BRFSS_USE$MARST <- as.factor(BRFSS_USE$MARITAL) 
levels(BRFSS_USE$MARST) <- c("married","divorced","widowed","separated","never married","living tgt")
BRFSS_USE$MARST <- fct_collapse(BRFSS_USE$MARST, never.married = c("living tgt","never married"))
fct_relevel(BRFSS_USE$MARST , "divorced", "married", "never.married","separated", "widowed")


#Race

is.na(BRFSS_USE$X_IMPRACE) <- which(BRFSS_USE$X_IMPRACE == 77)
is.na(BRFSS_USE$X_IMPRACE) <- which(BRFSS_USE$X_IMPRACE == 99)
na.exclude(BRFSS_USE$X_IMPRACE)
BRFSS_USE$RACE <- as.factor(BRFSS_USE$X_IMPRACE) #Have to change name in BRFSS
levels(BRFSS_USE$RACE) <- c("White","Black", "American Indian/Alaskan Native", "Asian", "Pacific","Other Race","Multiracial")
BRFSS_USE$RACE <- fct_collapse(BRFSS_USE$RACE, Asian = c("Asian", "Pacific"))
fct_relevel(BRFSS_USE$RACE, "American Indian/Alaskan Native", "Asian", "Black","Multiracial", "Other Race", "White")



#Hispanic
is.na(BRFSS_USE$X_HISPANC) <- which(BRFSS_USE$X_HISPANC == 9)
na.exclude(BRFSS_USE$X_HISPANC)
BRFSS_USE$X_HISPANC <- as.factor(BRFSS_USE$X_HISPANC)
levels(BRFSS_USE$X_HISPANC) <- c("Yes","No")

#Health Coverage

is.na(BRFSS_USE$HLTHPLN1) <- which(BRFSS_USE$HLTHPLN1 == 9)
is.na(BRFSS_USE$HLTHPLN1) <- which(BRFSS_USE$HLTHPLN1 == 7)
na.exclude(BRFSS_USE$HLTHPLN1)
BRFSS_USE$HLTHPLN1 <- as.factor(BRFSS_USE$HLTHPLN1)
levels(BRFSS_USE$HLTHPLN1) <- c("Yes","No")
fct_relevel(BRFSS_USE$HLTHPLN1, "No", "Yes")


#In Labor force

is.na(BRFSS_USE$EMPLOY1) <- which(BRFSS_USE$EMPLOY1 == 9)
na.exclude(BRFSS_USE$EMPLOY1)
BRFSS_USE$LABFORCE <- as.factor(BRFSS_USE$EMPLOY1)
levels(BRFSS_USE$LABFORCE) <- c("Employed wage","self-employed","out of work","out of work 1<", "homemaker", "student","retired","disabled")
BRFSS_USE$LABFORCE <- fct_collapse(BRFSS_USE$LABFORCE, Not.in.LF = c("out of work","homemaker","student","retired","disabled"))
BRFSS_USE$LABFORCE <- fct_collapse(BRFSS_USE$LABFORCE, in.LF = c("Employed wage","self-employed","out of work 1<"))
fct_relevel(BRFSS_USE$LABFORCE, "Not.in.LF", "in.LF")

#Education
is.na(BRFSS_USE$EDUCA) <- which(BRFSS_USE$EDUCA == 9)
na.exclude(BRFSS_USE$EDUCA)
BRFSS_USE$EDUCA <- as.factor(BRFSS_USE$EDUCA)
levels(BRFSS_USE$EDUCA) <- c("kinder","Grades 1 through 8","Grades 9 through 11","Grade 12 or GED", "College 1 year to 3 years", "College 4 years or more")
BRFSS_USE$EDUCA <- fct_collapse(BRFSS_USE$EDUCA, Elementary = c("kinder","Grades 1 through 8"))
BRFSS_USE$EDUCA <- fct_collapse(BRFSS_USE$EDUCA, SomeHS = c("Grades 9 through 11"))
BRFSS_USE$EDUCA <- fct_collapse(BRFSS_USE$EDUCA, HS = c("Grade 12 or GED"))
BRFSS_USE$EDUCA <- fct_collapse(BRFSS_USE$EDUCA, SomeC = c("College 1 year to 3 years"))
BRFSS_USE$EDUCA <- fct_collapse(BRFSS_USE$EDUCA, College = c("College 4 years or more"))

```


```{r}
#library(mice)
#init = mice(BRFSS_USE, maxit = 0) 
#meth = init$method
#predM = init$predictorMatrix
```



#Random forest for the likeability of a person in NYS being in NYC

```{r}
require("standardize")
set.seed(654321)
NN1 <- length(acs2017_ny$in_NYC)
#restrict_1 <- as.logical(round(runif(NN1,min=0,max=0.70))) # use fraction as training data
restrict_1 <- (runif(NN1) < 0.20) # use 10% as training data
summary(restrict_1)
dat_train1 <- subset(acs2017_ny, restrict_1)
dat_test1 <- subset(acs2017_ny, !restrict_1)
sobj1 <- standardize(in_NYC ~  INCOME2 + RENTHOM1 + RACE + X_HISPANC + EDUCA + LABFORCE + HLTHPLN1 + AGE + MARST + SEXVAR, dat_train1, family = binomial, na.action = na.pass)

s_dat_test1 <- predict(sobj1, dat_test1)
```

```{r}
require('randomForest')
set.seed(54321)
model_rf1 <- randomForest(in_NYC ~., data = sobj1$data, na.action = na.exclude)
print(model_rf1)
round(importance(model_rf1),2)
varImpPlot(model_rf1)
# look at confusion matrix for this too
pred_model1 <- predict(model_rf1, s_dat_test1)



```
#Should we also restric metropolitan area at this point? #here
Subset for people in NYS only *BRFSS 
```{r}
BRFSS.SUBSET.RULE <- (BRFSS_USE$X_STATE == 36) & (BRFSS_USE$X_METSTAT == 1)
BRFSS.SUBSET <- subset(BRFSS_USE,BRFSS.SUBSET.RULE) 

#attach(BRFSS.SUBSET)
```



```{r}
s_dat_test2 <- predict(sobj1, newdata = BRFSS.SUBSET)
BRFSS.SUBSET$in_NYC <- predict(model_rf1, newdata = s_dat_test2,)
summary(BRFSS.SUBSET$in_NYC)
```

Now, we do borough predictions 

```{r}
acs2017_ny$borough <- factor((in_Bronx + 2*in_Manhattan + 3*in_StatenI + 4*in_Brooklyn + 5*in_Queens), levels=c(1,2,3,4,5),labels = c("Bronx","Manhattan","Staten Island","Brooklyn","Queens"))

```

```{r}
require("standardize")
set.seed(54321)
NN <- length(acs2017_ny$borough)
restrict_2 <- as.logical(round(runif(NN,min=0,max=0.75))) # use fraction as training data
#restrict_1 <- (runif(NN) < 0.1) # use 10% as training data
summary(restrict_2)
dat_train2 <- subset(acs2017_ny, restrict_2)
dat_test2 <- subset(acs2017_ny, !restrict_2)
sobj2 <- standardize(borough ~ INCOME2 + RENTHOM1 + RACE + X_HISPANC + EDUCA + LABFORCE  + HLTHPLN1 + AGE + MARST, dat_train2, family = binomial)

s_dat_test3 <- predict(sobj2, dat_test2)
```

```{r}

require('randomForest')
set.seed(54321)
model_rf2 <- randomForest(borough ~., data = sobj2$data, na.action = na.omit)
print(model_rf2)
round(importance(model_rf2),2)
varImpPlot(model_rf2)
# look at confusion matrix for this too
pred_model2 <- predict(model_rf2,  s_dat_test3)
table(pred = pred_model2, true = dat_test2$borough)

```


```{r}
BRFSS.SUBSET.RULE2 <- (BRFSS.SUBSET$in_NYC == "Yes")
BRFSS.SUBSET2 <- subset(BRFSS.SUBSET,BRFSS.SUBSET.RULE2) 
```


```{r}
s_dat_test3 <- predict(sobj2, BRFSS.SUBSET2)
BRFSS.SUBSET2$borough <- predict(model_rf2, s_dat_test3, na.action=na.exclude)
summary(BRFSS.SUBSET2$borough)
```
Recoding for the fruit variables 

#Fruits daily 
```{r}
#NAs 
is.na(BRFSS.SUBSET2$FRUIT2) <- which(BRFSS.SUBSET2$FRUIT2 == 777)
is.na(BRFSS.SUBSET2$FRUIT2) <- which(BRFSS.SUBSET2$FRUIT2 == 999)
#Creates Dummy
BRFSS.SUBSET2$Daily.fruit <- ifelse((BRFSS.SUBSET2$FRUIT2 >= 101) & (BRFSS.SUBSET2$FRUIT2 <= 199), 1, 0)

BRFSS.SUBSET2$Daily.fruit  <- as.factor(BRFSS.SUBSET2$Daily.fruit)
levels(BRFSS.SUBSET2$Daily.fruit) <- c("Yes", "No")

```

#Veggies daily 
```{r}
#NAs 
is.na(BRFSS.SUBSET2$FVGREEN1) <- which(BRFSS.SUBSET2$FVGREEN1== 777)
is.na(BRFSS.SUBSET2$FVGREEN1) <- which(BRFSS.SUBSET2$FVGREEN1 == 999)
#Creates Dummy
BRFSS.SUBSET2$Daily.greens <- ifelse((BRFSS.SUBSET2$FVGREEN1 >= 101) & (BRFSS.SUBSET2$FVGREEN1 <= 199), 1, 0)

BRFSS.SUBSET2$Daily.greens  <- as.factor(BRFSS.SUBSET2$Daily.greens)
levels(BRFSS.SUBSET2$Daily.greens) <- c("Yes", "No")
```


```{r}
Model.Logit <- glm(Daily.fruit ~ borough,
            family = binomial, data = BRFSS.SUBSET2)

summary(Model.Logit)
```

```{r}
Model.Logit2 <- glm(Daily.greens ~ borough,
            family = binomial, data = BRFSS.SUBSET2)

summary(Model.Logit2)
require("effects")
plot(allEffects(Model.Logit4))
```

```{r}
BRFSS.SUBSET2$Healthy.ppl <- ifelse((BRFSS.SUBSET2$Daily.fruit == "Yes") | (BRFSS.SUBSET2$Daily.greens == "Yes"), 1, 0)
```

```{r}
Model.Logit3 <- glm(Healthy.ppl ~ borough,
            family = binomial, data = BRFSS.SUBSET2)

summary(Model.Logit3)

```

```{r}
BRFSS.SUBSET2$Healthy.ppl  <- as.factor(BRFSS.SUBSET2$Healthy.ppl)
levels(BRFSS.SUBSET2$Healthy.ppl) <- c("Yes", "No")

Model.Logit4 <- glm(Healthy.ppl ~ INCOME2,
            family = binomial, data = BRFSS.SUBSET2)

summary(Model.Logit4)
require("effects")
plot(allEffects(Model.Logit4))
```

```{r}
Model.Logit4 <- glm(Healthy.ppl ~ borough,
            family = binomial, data = BRFSS.SUBSET2)

summary(Model.Logit4)
require("effects")
plot(allEffects(Model.Logit4))

```






